Master thesis

### 2017-07-13 02-Introduction.Rmd
Thesis v1

But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging. We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better syslvicultural practives in order to reach sustainability.

@@

Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity. We assumed that when a disturbance event happen, due to a higher productivity, foerst with an increased diversity with recover quickly and thus be more resilient. This theoritical expectation stand on two processes: complementarity and diversity effects [@Loreau2001]. If we take a species pools with different productivity in monoculture. Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressource partitionning and niche differentiation. Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage. Complementarity effect is the addition of the better ressource acquisition and the facilitation. Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection. And if competitive species are more productive, they will increase assemblage overall productivity. This ressources preemption by more competitive species is the selection effect.

Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests. One of the few studies was realized by @Chisholm2013 and has shwon a positive significative relationship between species richness and wood productivity on a worlwide forest network. But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.

In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience, more particularly tropical forest carbon resilience. We used a simulation approach using TROLL model to assess long term processes.

>>>>

But several authors argue that selective logging represents a major threat to biodiversity [@Carreno-Rocabado2012; @DeAvila2015; @Gibson2013; @Martin2015; @Zimmerman2012], challenging the sustainable definition from current selective logging. We consequently need to assess both short and long term impacts of selective logging on tropical forest ecosystems to implement better sylvicultural practices in order to reach sustainability.

@@

Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity. We assumed that when a disturbance event happen, due to a higher productivity, forest with an increased diversity will recover quicker and thus be more resilient. This theoritical expectation stand on two processes: complementarity and diversity effects [@Loreau2001]. If we take a species pools with different productivity in monoculture. Their assemblage will allow theim an overall higher productivity due to a better ressources acquisition through ressource partitionning and niche differentiation. Additionally, some species will individually have an higher productivity in the assemblage than in monoculture due to facilitation from other species present in the assemblage. Complementarity effect is the addition of the better ressource acquisition and the facilitation. Now if we look at the evolution of the species assemblage over time, more competitive species will progressively dominate through competitive selection. And if competitive species are more productive, they will increase assemblage overall productivity. This ressources preemption by more competitive species is the selection effect.

Generally, positive relationship between biodiversity and productivity were emprically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of tropical forests. One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worlwide forest network. But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.

In the present study, we focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience. We used a simulation approach using TROLL model to assess long term processes.

### 2017-07-13 03-Model.Rmd
Thesis v1

```{r TROLLtree, echo=FALSE, fig.cap='Individuals tree inside TROLL explicit spatial grid from @Li. Tree geometry (crown radius CR, crown depth CD, height h, diameter at breast height dbh) is updated at each timestep following allometric relationship with assimilated carbon allocated to growth. Each tree is flagged with a species label linking to its species-specific attributes. Light is compputed explicitly at each timestep for each voxel.'}
knitr::include_graphics('images/TROLLtree.png', dpi = NA)
```

@@

In french guiana context sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting. Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber. Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).

@@

Tree harvesting roads are split in three classes: truck roads, main tractor track, and secondary track. Because TROLL assumes a flat environment, the main track is opened starting from the middle of one side of the simulated forest and untill it reaches the center with a width of 6 meters. In most cases, secondary tracks are opened once trees have been designated and the geolocation taken at a maximum distance of 30 meters from designated trees (Laurent Descroix, ONF, personnal communication). To simulate secondary tracks, TROLL uses a loads map, measuring every trees at a distance of 30 meters for each pixel, and  a track proximity maps of the closest existing track. Next, the model select the pixel with the highest load and closest track, find the closest existing track and join it by removing tree in the way with a width of 5 meters. The operations are repeated untill no felt trees are left.

@@

$Death$ of a tree follows a $\mathcal{Bernoulli}$ law of probability $P(Death)$. The odds for a tree to die are calculated with the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$. The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap. The probability for a tree to die $P(Death)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd.

>>>>

```{r TROLLtree, echo=FALSE, fig.cap='Individuals tree inside TROLL explicit spatial grid from @Li. Tree geometry (crown radius CR, crown depth CD, height h, diameter at breast height dbh) is updated at each timestep following allometric relationship with assimilated carbon allocated to growth. Each tree is flagged with a species label linking to its species-specific attributes. Light is computed explicitly at each timestep for each voxel.'}
knitr::include_graphics('images/TROLLtree.png', dpi = NA)
```

@@

In french guiana context, sylviculture can be narrow to selective logging, which can be split in two steps: selection and harvesting. Selection encompass choice of the havrestable area, tree designation by the forest office, tree selection by the harvester, and removal off tree probbed as rotten by the lumber. Harvesting encompass tree felling, tracks opening, and long term damages (simplified in gap damages in current TROLL implementation).

@@

Tree harvesting roads are split in three classes: truck roads, main tractor track, and secondary track. Because TROLL assumes a flat environment, the main track is opened starting from the midle of one side of the simulated forest and untill it reaches the center with a width of 6 meters. In most cases, secondary tracks are opened once trees have been designated and the geolocation taken at a maximum distance of 30 meters from designated trees (Laurent Descroix, ONF, personnal communication). To simulate secondary tracks, TROLL uses a loads map, measuring every trees at a distance of 30 meters for each pixel, and  a track proximity maps of the closest existing track. Next, the model select the pixel with the highest load and closest track, find the closest existing track and join it by removing tree in the way with a width of 5 meters. The operations are repeated untill no felt trees are left.

@@

$Death$ of a tree follows a $\mathcal{B}ernoulli$ law of probability $P(Death)$. The odds for a tree to die are calculated with the sum of the natural tree death odd $\theta$ and a perturbation index $\beta*e^{\alpha*d_{gaps}}$. The perturbation index depend on the distance $d_{gaps}$ of the tree $i$ to the closest logging gap. The probability for a tree to die $P(Death)$ is finally calculated by taking the inverse logit $logit^{-1}$ of the odd.

### 2017-07-13 04-MnMs.Rmd
Introduction v2
Thesis v1

@Li already assessed TROLL model sensitivity to several parameters ($k$ see \@ref(eq:PPFD), $\phi$ see \@ref(eq:LCP), $g1$ see \@ref(eq:gs), $f_{wood}$ see \@ref(eq:DeltaV), $f_{canopy}$ see \@ref(eq:DeltaLA) and $m$ see \@ref(eq:db)) which they assumed having a key role in model functioning. On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversity. Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities. We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assumed it as one of the main factors of tree recruitments after disturbance within simulations.

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory. We assumed maturity of the forest after 500 years of regeneration (Maréchaux & Chave) and computed simulation 100 years after a disturbance event with 40% loss of basal area. Due to computer limitations we did not run replicate (besides it should be necessary to reduce simulation stochasticity). To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values. Ecosystem outputs outside of the range of the control replicates values are significantly influenced by the studied parameter.

@@

In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time. Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation. Biodiversity was integrated with two components of its components: taxonomic and functional diversities. We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species). Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean. Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess theim before diturbance. Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself. This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]). Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance. We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.

@@

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic. Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the the ecosystem metric values from the control $X_C(t)$ . Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics. $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest dynamic (AGB, BA, N, N10, and N30) and forest production (GPP, NPP, Rday, and Rnight). We then used cummulative integral over time of eulcidean distance to equilibrium to assess simulations resilience.

>>>>

@Li already assessed TROLL model sensitivity to several parameters ($k$ see \@ref(eq:PPFD), $\phi$ see \@ref(eq:LCP), $g1$ see \@ref(eq:gs), $f_{wood}$ see \@ref(eq:DeltaV), $f_{canopy}$ see \@ref(eq:DeltaLA) and $m$ see \@ref(eq:db)) which they assumed having a key role in model functioning. On the other hand, we decided to use TROLL to study resistance and resilience of ecosystem face to disturbance, highlighting the role of biodiversity. Consequently we particulrarly needed to assess the importance of functional traits to further better control and evaluate functional diversities. We also needed to assess the sensitivity of TROLL model to the seed rain constant ($n_{ext}$, see \@ref(eq:next)) because we assumed it was one of the main factors of tree recruitments after disturbance within simulations.

@@

Simulations were conducted on Intel Xeon(R) with 32 CPUs of 2.00GHz and 188.9 GB of memory. We assumed maturity of the forest after 500 years of regeneration @Li and computed simulation 100 years after a disturbance event with 40% loss of basal area. Due to computer limitations we did not run replicates (besides it should be necessary to reduce simulation stochasticity). To assess ecosystem outputs sensitivity to studied parameters, we compared it to 100 replicates of control simulations with all parameters set to default values. Ecosystem outputs outside of the range of the control replicates values are significantly influenced by the studied parameter.

@@

In order to assess the role of biodiversity in ecosystem answer to both disturbance and sylviculture, we needed to create a space of experiments encompassing both variation of disturbance, biodiversity and time. Disturbance was represented by percentage of basal area loss (0%, 25%, 50% and 75%), or as a selective logging simulation using default parameters. Biodiversity was integrated with two of its components: taxonomic and functional diversities. We used species richness $SR$ to represents taxonomic diversity (5, 25, and 125 species). Functional diversity can be related to numerous components, and @Borgy2017 argued for 5: richness, divergence, regularity, overlap and mean. Because mature forest were created from a bare soil with TROLL simulations, we could not control a priori divergence, regularity and overlap but only assess theim before diturbance. Consequently, we focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we selected 20 communities with growing convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. Effectivelly, we did not wanted drastic change in community means that could have more effect than functional richness itself. This design of experiments resulted in 60 communities ($5~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Li]). Figure \@ref(fig:DOE) presents the design of experiment for communities biodiversity after the mature forest were simulated, and thus before disturbance. We obtained a broad range of both functionl dispersion $FDis$ and aboveground biomass $AGB$ for simulated forest ecosystems before disturbance.

@@

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of tropical forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ can not be reduced to a specific time if the equilibrium is dynamic. Consequently, to encompass undisturbed ecosystem variations throught time, we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ . Thus, the value of resilience $R(t)$ is normalized for all simulations and metrics. $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Consequently we can calculate an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for the two functions described above: forest dynamic (AGB, BA, N, N10, and N30) and forest production (GPP, NPP, Rday, and Rnight). We then used cummulative integral over time of eulcidean distance to equilibrium to assess simulations resilience.

### 2017-07-13 05-Results.Rmd
Thesis v1

Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)). On the other hand, few functional traits influenced final ecosystem structure. Only specific maximum diameter **dmax** add higher diversity for greater orders implying better eveness in species distributions. Regarding functional composition, traits fixed to mean did not change other functional traits density distribution. Moreover, seedrain did not seem to affect aboveground biomass and final ecosystem height and diameter structure. Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect. Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm diameter at breast height after ecosystem resilience to disturbance (approximately 50 years). Lower seedrain than default decreased basal area over time. In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species. See [Appendix 4: Sensitivity analysis] for further details.

@@

The ranking was stable over time for the 240 simulations. So we used the cummulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience. We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)). We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years ($Ieq_{600}$). In addition, functional eveness was complementary reducing $Ieq_{600}$. Finally species richness was not directly link to $Ieq_{600}$. Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 5: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional eveness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\neveness\n(FEve)')
```

>>>>

Most of functional traits had a significant long term influence on ecosystem outputs (figure \@ref(fig:ftVar)). On the other hand, few functional traits influenced final ecosystem structure. Only specific maximum diameter **dmax** add higher diversity for greater orders implying better evenness in species distributions. Regarding functional composition, traits fixed to mean did not change other functional traits density distribution. Moreover, seedrain did not seem to affect aboveground biomass and final ecosystem height and diameter structure. Seedrain constant fixed to 2 or 20 seed per hectare seemed to have a similar effect. Lower seedrain implied faster decrease of stem above 10 cm dbh and higher number of stem above 30 cm diameter at breast height after ecosystem resilience to disturbance (approximately 50 years). Lower seedrain than default decreased basal area over time. In addition, lower seedrain than default decreased equitability by increasing abundance of abundant species and decreasing abundance of less abundant species. See [Appendix 4: Sensitivity analysis] for further details.

@@

The ranking was stable over time for the 240 simulations. So we used the cummulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience. We compared cummulative integral after 600 years to communities taxonomic and functional diversity for each leavel of disturbance (see Figure \@ref(fig:gIeq)). We found that increased functional diversity [FDiv, @villeger_new_2008] was reducing cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years ($Ieq_{600}$). In addition, functional evenness was complementary reducing $Ieq_{600}$. Finally species richness was not directly link to $Ieq_{600}$. Effectivelly, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels and other ecosystem functions (forest production and forest diversity, see [Appendix 5: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest dynamic equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008].', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest dynamic equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)')
```

### 2017-07-13 06-Discussion.Rmd
Thesis v1

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting are not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

@@

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and eveness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional eveness. Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of eveness  confirm the review of @Zhang2012 who highlighted the role of eveness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses is not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the major facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shwon by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the case. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two rotation in order to maintain both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience and that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience, we were able to show, of tropical forest hyperdiverse systems and let hope for a future sustainable selective logging in tropical forests.

>>>>

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting is not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

@@

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and evenness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional evenness. Effectively high functional diversity needs evenness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of evenness  confirm the review of @Zhang2012 who highlighted the role of evenness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses is not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

@@

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the main facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shwon by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the cases. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two cutting cycles in order to maintain both productive ecosystem and sustainable management.

@@

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience. Additionally we found that complementarity between species was insuring forest recovery in forest succession start with facilitation, before more productive species dominate the forest and insure forest recovery. Our results advocate for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience of tropical forest hyperdiverse systems, we were able to show, and let hope for a future sustainable selective logging in tropical forests.

### 2017-07-13 08-Appendix2.Rmd
Results: Disturbance & Appendix4
Thesis v1

We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)). We kept dataset given by GLOPNET as origin dataset for observations. Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$). We measured variable importance in functionnal traits to explain leaf lifespan with out-of bag method applied on a random forest. Then, we used a béayesian apporach to test different models with growing level of complexity. We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$). We finally tested the new allometry obtained with the selected model with TROLL simulations.

@@

```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values. Leaf lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp. Nmass and wsg) is not constant and depend on the closest point value for right (resp. center and left) graph.', cahce=TRUE, fig.height=12, fig.width=12}

@@

Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early edath of individuals inside simulations. For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.

>>>>

We compiled functional traits from GLOPNET [@wright_worldwide_2004], TRY [@Kattge2011], and DRYAD [@chave_towards_2009] databases (see \@ref(tab:A2traits)). We kept dataset given by GLOPNET as origin dataset for observations. Dataset defined as origin corresponded to leaf lifespan ($LL$) and most of the time to leaf mass per area ($LMA$) and leaf nitrogen content per leaf dry mass ($Nmass$). We measured variable importance in functionnal traits to explain leaf lifespan with an out-of the bag method applied on a random forest. Then, we used a bayesian apporach to test different models with growing level of complexity. We retained the model with the best tradeoff between model complexity (number of parameters $K$), convergence, likelihood, and prediction quality (root mean sqaure error of prediciton $RMSEP$). We finally tested the new allometry obtained with the selected model with TROLL simulations.

@@

```{r A2LLpred, echo=FALSE, fig.cap='Leaflifespan predictions for the selected model with leaf mass per area (LMA), leaf nitrogen content (Nmass), wood specific gravity (wsg) and predicted versus observed values. Leaf lifespan (LL) is predicted with model M10 fit. Leaf mass per area (LMA) and leaf nitrogen content (Nmass), and wood specific gravity (wsg) are taken in a composite dataset of GLOPNET, TRY and DRYAD  datasets.Warning LMA (resp. Nmass and wsg) is not constant and depend on the closest point value for right (resp. center and left) graph.', cache=TRUE, fig.height=12, fig.width=12}

@@

Simulations are validating that this new allometry resolve the issue of unrealistically low aboveground biomass for low LMA species due to an early death of individuals inside simulations. For instance with this allometry Symphonia sp 1 (a low LMA species) is now reaching a realistic aboveground biomass above 400 $tonC.ha^{-1}$ and realistic diameter and age distribution inside the final population.

### 2017-07-13 09-Appendix3.Rmd
Thesis v1

In order to simulate sylviculture with TROLL we need to implement a new sylviculture module inside TROLL model code. A first litterature review was completed by an interview with Laurent Descroix of the Office Nationale des Forêts. We discovered that rotten trees were not random and seemed to depend both on tree species and diameter. This document presents modelling of relation between rotten trees and their species and diameter.

In fact we have two different questions:

@@

$$V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$

@@

```{r A3models, echo=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/rotten_models"
mtab <- data.frame(
  M = c('$M_{s,p}$', '$N_{s,p} + L_{s,p}$'),
  Model = c(
    "${P_{rotten}}_n \\sim \\mathcal{B}(inv_{logit}(\\beta_0 + \\beta_1*dbh_n + {\\beta_2}_p + {\\beta_3}_s))$",
    "$Volume_{of~wood} \\sim \\mathcal{logN} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
    )
  )
mtab <- apply(mtab, 2, as.character)
knitr::kable(mtab, caption = 'Models summary.', format = 'pandoc')
row.names(mtab) <- c('Mps', 'Nps')
```

>>>>

In order to simulate sylviculture with TROLL we needed to implement a new sylviculture module inside TROLL model code. A first litterature review was completed by an interview with Laurent Descroix of the Office Nationale des Forêts. We discovered that rotten trees were not random and seemed to depend both on tree species and diameter. This document presents modelling of relation between rotten trees and their species and diameter.

In fact we have two different objectives:

@@

$$V_f \sim log\mathcal{N}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$

@@

```{r A3models, echo=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/rotten_models"
mtab <- data.frame(
  M = c('$M_{s,p}$', '$N_{s,p} + L_{s,p}$'),
  Model = c(
    "${P_{rotten}}_n \\sim \\mathcal{B}(inv_{logit}(\\beta_0 + \\beta_1*dbh_n + {\\beta_2}_p + {\\beta_3}_s))$",
    "$Volume_{of~wood} \\sim log\\mathcal{N} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
    )
  )
mtab <- apply(mtab, 2, as.character)
knitr::kable(mtab, caption = 'Models summary.', format = 'pandoc')
row.names(mtab) <- c('Mps', 'Nps')
```
